---
title: "Example1RunFile"
author: "CutPosteriorAuthors"
date: "2024-05-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load required packages
```{r}
library(foreach)
library(doParallel)
library(pracma)
library(copula)
library(extraDistr)
library(mvtnorm)
library(numDeriv)
library(Matrix)
```

#Load customised functions
```{r}
source("NewMCMCcodesExample1Revised.R")
source("NewVBcodesExample1Revised.R")
```

#Select training sample size
```{r}
N <- 1000
#N <- 500
#N <- 100
```

#Set true copula density
```{r}
d <- 2
tau.true <- 0.7
cc <- tCopula(param = sin(0.5*pi*tau.true) ,dim = d, df=1)
```

#Set true parameter values
```{r}
true.val <- c(1,1,7,3,tau.true)
```

#MCMC inference with correctly specified marginals and copula 
```{r}
theta.correct <- matrix(0,nrow=500,ncol=5)
theta.var.correct <- matrix(0,nrow=500,ncol=5)
lower.correct <- matrix(0,nrow=500,ncol=5)
upper.correct <- matrix(0,nrow=500,ncol=5)
for(seed.num in 1:500){
  
  set.seed(seed.num)
  U <- rCopula(N, copula = cc)
  y1 <- qlnorm(U[,1], meanlog = 1, sdlog = 1)
  y2 <- qgamma(U[,2], shape=7, rate = 3)
  MLE <-univariateML::mlgamma(y2)
  MCMCobj.correct <- NewMCMCfull.LogNormal.Gamma.T(y = cbind(y1,y2), mu.init = mean(log(y1)), v.init = var(log(y1)), a.init = log(MLE[1]), b.init = log(MLE[2]), omega.init = max(logit(2/pi*asin(cor(y1,y2))),0.05), B = 15000 )
  theta.correct[seed.num,] <- c(mean(MCMCobj.correct$mu.vec[5001:15000]), mean(exp(MCMCobj.correct$v.vec[5001:15000])), mean(exp(MCMCobj.correct$a.vec[5001:15000])), mean(exp(MCMCobj.correct$b.vec[5001:15000])), mean(MCMCobj.correct$tau.vec[5001:15000]))
  theta.var.correct[seed.num,] <- c(var(MCMCobj.correct$mu.vec[5001:15000]), var(exp(MCMCobj.correct$v.vec[5001:15000])), var(exp(MCMCobj.correct$a.vec[5001:15000])), var(exp(MCMCobj.correct$b.vec[5001:15000])), var(MCMCobj.correct$tau.vec[5001:15000]))
  lower.correct[seed.num,] <- c(quantile(MCMCobj.correct$mu.vec[5001:15000], 0.025), quantile(exp(MCMCobj.correct$v.vec[5001:15000]), 0.025), quantile(exp(MCMCobj.correct$a.vec[5001:15000]), 0.025), quantile(exp(MCMCobj.correct$b.vec[5001:15000]), 0.025), quantile(MCMCobj.correct$tau.vec[5001:15000], 0.025))
  upper.correct[seed.num,] <- c(quantile(MCMCobj.correct$mu.vec[5001:15000], 0.975), quantile(exp(MCMCobj.correct$v.vec[5001:15000]), 0.975), quantile(exp(MCMCobj.correct$a.vec[5001:15000]), 0.975), quantile(exp(MCMCobj.correct$b.vec[5001:15000]), 0.975), quantile(MCMCobj.correct$tau.vec[5001:15000], 0.975))
  
  cat("Completed seed number:", seed.num, "\n")
  
}
Bias.correct <- rowMeans(apply(theta.correct,1,function(s){ s- true.val }))
RMSE.correct <- sqrt(colMeans(t(apply(theta.correct,1,function(s){ (s - true.val)^2 })) ))
Coverage.correct <- rowMeans(apply(lower.correct,1,function(s){ s <= true.val })*apply(upper.correct,1,function(s){ s >= true.val }))
KL.marginal1.correct <- 0.5*( (theta.correct[,2]/1) - 1 + (theta.correct[,1] - 1)^2/1 + log(1/theta.correct[,2]) )
KL.marginal2.correct <- rep(0,500)
KS.cop.vec <- rep(0,500)
u1.values <- seq(0.01,0.99,0.01)
u2.values <- u1.values
for(rs in 1:500){
  
  KL.marginal2.correct[rs] <- KL.GAMMA(shape1 = theta.correct[rs,3], rate1 = theta.correct[rs,4], shape2 = true.val[3], rate2 = true.val[4])
  KLGumbelCop <- function(u1, u2){
    
    a1 <- LogPDFTCop(F1 = u1, F2 = u2, tau = theta.correct[rs,5])
    a2 <- LogPDFTCop(F1 = u1, F2 = u2, tau = true.val[5])
    return(exp(a1)*(a1-a2))
    
  }
  grid.temp <- outer(u1.values,u2.values, FUN=KLGumbelCop)
  KS.cop.vec[rs] <- max(0,0.25*0.01^2*(sum(grid.temp[1:98,1:98]) + sum(grid.temp[1:98,2:99]) + sum(grid.temp[2:99,1:98]) + sum(grid.temp[2:99,2:99])))
  
}
Bias.correct
RMSE.correct
Coverage.correct
c(mean(KL.marginal1.correct),mean(KL.marginal2.correct),mean(KS.cop.vec))
```

#MCMC inference with correctly specified marginals and misspecified copula
```{r}
theta.uncut <- matrix(0,nrow=500,ncol=5)
theta.var.uncut <- matrix(0,nrow=500,ncol=5)
lower.uncut <- matrix(0,nrow=500,ncol=5)
upper.uncut <- matrix(0,nrow=500,ncol=5)
for(seed.num in 1:500){
  
  set.seed(seed.num)
  U <- rCopula(N, copula = cc)
  y1 <- qlnorm(U[,1], meanlog = 1, sdlog = 1)
  y2 <- qgamma(U[,2], shape=7, rate = 3)
  MLE <-univariateML::mlgamma(y2)
  
  MCMCobj.uncut <- NewMCMCfull.LogNormal.Gamma.Gumbel(y = cbind(y1,y2), mu.init = mean(log(y1)), v.init = var(log(y1)), a.init = log(MLE[1]), b.init = log(MLE[2]), omega.init = max(logit(2/pi*asin(cor(y1,y2))),0.05), B = 15000 )
  theta.uncut[seed.num,] <- c(mean(MCMCobj.uncut$mu.vec[5001:15000]), mean(exp(MCMCobj.uncut$v.vec[5001:15000])), mean(exp(MCMCobj.uncut$a.vec[5001:15000])), mean(exp(MCMCobj.uncut$b.vec[5001:15000])), mean(MCMCobj.uncut$tau.vec[5001:15000]))
  theta.var.uncut[seed.num,] <- c(var(MCMCobj.uncut$mu.vec[5001:15000]), var(exp(MCMCobj.uncut$v.vec[5001:15000])), var(exp(MCMCobj.uncut$a.vec[5001:15000])), var(exp(MCMCobj.uncut$b.vec[5001:15000])), var(MCMCobj.uncut$tau.vec[5001:15000]))
  lower.uncut[seed.num,] <- c(quantile(MCMCobj.uncut$mu.vec[5001:15000], 0.025), quantile(exp(MCMCobj.uncut$v.vec[5001:15000]), 0.025), quantile(exp(MCMCobj.uncut$a.vec[5001:15000]), 0.025), quantile(exp(MCMCobj.uncut$b.vec[5001:15000]), 0.025), quantile(MCMCobj.uncut$tau.vec[5001:15000], 0.025))
  upper.uncut[seed.num,] <- c(quantile(MCMCobj.uncut$mu.vec[5001:15000], 0.975), quantile(exp(MCMCobj.uncut$v.vec[5001:15000]), 0.975), quantile(exp(MCMCobj.uncut$a.vec[5001:15000]), 0.975), quantile(exp(MCMCobj.uncut$b.vec[5001:15000]), 0.975), quantile(MCMCobj.uncut$tau.vec[5001:15000], 0.975))
  
}
Bias.uncut <- rowMeans(apply(theta.uncut,1,function(s){ s- true.val }))
RMSE.uncut <- sqrt(colMeans(t(apply(theta.uncut,1,function(s){ (s - true.val)^2 })) ))
Coverage.uncut <- rowMeans(apply(lower.uncut,1,function(s){ s <= true.val })*apply(upper.uncut,1,function(s){ s >= true.val }))
KL.marginal1.uncut <- 0.5*log(1) - 0.5*log(theta.uncut[,2]) + 0.5*(theta.uncut[,2]/1) + 0.5*(theta.uncut[,1]-1)^2 - 0.5
KL.marginal2.uncut <- rep(0,500)
KS.cop.vec <- rep(0,500)
u1.values <- seq(0.01,0.99,0.01)
u2.values <- u1.values
for(rs in 1:500){
  
  KL.marginal2.uncut[rs] <- KL.GAMMA(shape1 = theta.uncut[rs,3], rate1 = theta.uncut[rs,4], shape2 = true.val[3], rate2 = true.val[4])
  KLGumbelCop <- function(u1, u2){
    
    a1 <- LogPDFGumbelCop(F1 = u1, F2 = u2, tau = theta.uncut[rs,5])
    a2 <- LogPDFTCop(F1 = u1, F2 = u2, tau = true.val[5])
    return(exp(a1)*(a1-a2))
    
  }
  grid.temp <- outer(u1.values,u2.values, FUN=KLGumbelCop)
  KS.cop.vec[rs] <- max(0,0.25*0.01^2*(sum(grid.temp[1:98,1:98]) + sum(grid.temp[1:98,2:99]) + sum(grid.temp[2:99,1:98]) + sum(grid.temp[2:99,2:99])))
  
}
Bias.uncut
RMSE.uncut
Coverage.uncut
c(mean(KL.marginal1.uncut),mean(KL.marginal2.uncut),mean(KS.cop.vec))
```

#MCMC Cut I posterior inference with misspecified copula
```{r}
theta.cut1 <- matrix(0,nrow=500,ncol=5)
theta.var.cut1 <- matrix(0,nrow=500,ncol=5)
lower.cut1 <- matrix(0,nrow=500,ncol=5)
upper.cut1 <- matrix(0,nrow=500,ncol=5)
for(seed.num in 1:500){
  
  set.seed(seed.num)
  U <- rCopula(N, copula = cc)
  y1 <- qlnorm(U[,1], meanlog = 1, sdlog = 1)
  y2 <- qgamma(U[,2], shape=7, rate = 3)
  MLE <-univariateML::mlgamma(y2)
  MCMCobj.marginal1.cut1 <- NewMCMCcutMarginal.LogNormal(y = y1, mu.init = mean(log(y1)), v.init = var(log(y1)), B = 15000)
  MCMCobj.marginal2.cut1 <- NewMCMCcutMarginal.Gamma(y = y2, a.init = log(MLE[1]), b.init = log(MLE[2]), B = 15000)
  # 
  omega.guess <- max(logit(2/pi*asin(cor(y1,y2))),0.05)
  cores=detectCores()
  cl <- makeCluster(cores[1]-1) #not to overload your computer
  registerDoParallel(cl)
  storeMATRIX <-  foreach(l=1:10000, .combine=cbind) %dopar% {

    library(copula)
    library(pracma)
    library(mvtnorm)
    MCMCobj.cop.cut1 <- NewMCMCcut1.LogNormal.Gamma.Gumbel(y = cbind(y1,y2), mu.val = MCMCobj.marginal1.cut1$mu.vec[5000+l], v.val = MCMCobj.marginal1.cut1$v.vec[5000+l], a.val = MCMCobj.marginal2.cut1$a.vec[5000+l], b.val = MCMCobj.marginal2.cut1$b.vec[5000+l], omega.init = omega.guess, B = 500)
    c(MCMCobj.cop.cut1$tau.vec[500])

  }
   stopCluster(cl)
  theta.cut1[seed.num,] <- c(mean(MCMCobj.marginal1.cut1$mu.vec[5001:15000]), mean(exp(MCMCobj.marginal1.cut1$v.vec[5001:15000])), mean(exp(MCMCobj.marginal2.cut1$a.vec[5001:15000])), mean(exp(MCMCobj.marginal2.cut1$b.vec[5001:15000])), mean(storeMATRIX[1,]))
  theta.var.cut1[seed.num,] <- c(var(MCMCobj.marginal1.cut1$mu.vec[5001:15000]), var(exp(MCMCobj.marginal1.cut1$v.vec[5001:15000])), var(exp(MCMCobj.marginal2.cut1$a.vec[5001:15000])), var(exp(MCMCobj.marginal2.cut1$b.vec[5001:15000])), var(storeMATRIX[1,]))
  lower.cut1[seed.num,] <- c(quantile(MCMCobj.marginal1.cut1$mu.vec[5001:15000],0.025), quantile(exp(MCMCobj.marginal1.cut1$v.vec[5001:15000]), 0.025), quantile(exp(MCMCobj.marginal2.cut1$a.vec[5001:15000]), 0.025), quantile(exp(MCMCobj.marginal2.cut1$b.vec[5001:15000]), 0.025), quantile(storeMATRIX[1,], 0.025))
  upper.cut1[seed.num,] <- c(quantile(MCMCobj.marginal1.cut1$mu.vec[5001:15000],0.975), quantile(exp(MCMCobj.marginal1.cut1$v.vec[5001:15000]), 0.975), quantile(exp(MCMCobj.marginal2.cut1$a.vec[5001:15000]), 0.975), quantile(exp(MCMCobj.marginal2.cut1$b.vec[5001:15000]), 0.975), quantile(storeMATRIX[1,], 0.975))
  cat("Completed seed number:", seed.num, "\n")
    
}
Bias.cut1 <- rowMeans(apply(theta.cut1,1,function(s){ s- true.val }))
RMSE.cut1 <- sqrt(colMeans(t(apply(theta.cut1,1,function(s){ (s - true.val)^2 })) ))
Coverage.cut1 <- rowMeans(apply(lower.cut1[,],1,function(s){ s <= true.val })*apply(upper.cut1,1,function(s){ s >= true.val }))
KL.marginal1.cut1 <- 0.5*log(1) - 0.5*log(theta.cut1[,2]) + 0.5*(theta.cut1[,2]/1) + 0.5*(theta.cut1[,1]-1)^2 - 0.5
KL.marginal2.cut1 <- rep(0,500)
KS.cop.vec <- rep(0,500)
u1.values <- seq(0.01,0.99,0.01)
u2.values <- u1.values
for(rs in 1:500){
  
  KL.marginal2.cut1[rs] <- KL.GAMMA(shape1 = theta.cut1[rs,3], rate1 = theta.cut1[rs,4], shape2 = true.val[3], rate2 = true.val[4])
  KLGumbelCop <- function(u1, u2){
    
    a1 <- LogPDFGumbelCop(F1 = u1, F2 = u2, tau = theta.cut1[rs,5])
    a2 <- LogPDFTCop(F1 = u1, F2 = u2, tau = true.val[5])
    return(exp(a1)*(a1-a2))
    
  }
  grid.temp <- outer(u1.values,u2.values, FUN=KLGumbelCop)
  KS.cop.vec[rs] <- max(0,0.25*0.01^2*(sum(grid.temp[1:98,1:98]) + sum(grid.temp[1:98,2:99]) + sum(grid.temp[2:99,1:98]) + sum(grid.temp[2:99,2:99])))
  
}
Bias.cut1
RMSE.cut1
Coverage.cut1
c(mean(KL.marginal1.cut1),mean(KL.marginal2.cut1),mean(KS.cop.vec))
```

#MCMC Cut I posterior inference with correctly specified copula
```{r}
theta.correct.cut1 <- matrix(0,nrow=500,ncol=5)
theta.var.correct.cut1 <- matrix(0,nrow=500,ncol=5)
lower.correct.cut1 <- matrix(0,nrow=500,ncol=5)
upper.correct.cut1 <- matrix(0,nrow=500,ncol=5)
for(seed.num in 1:500){
  
  set.seed(seed.num)
  U <- rCopula(N, copula = cc)
  y1 <- qlnorm(U[,1], meanlog = 1, sdlog = 1)
  y2 <- qgamma(U[,2], shape=7, rate = 3)
  MLE <-univariateML::mlgamma(y2)
  MCMCobj.marginal1.correct.cut1 <- NewMCMCcutMarginal.LogNormal(y = y1, mu.init = mean(log(y1)), v.init = var(log(y1)), B = 15000)
  MCMCobj.marginal2.correct.cut1 <- NewMCMCcutMarginal.Gamma(y = y2, a.init = log(MLE[1]), b.init = log(MLE[2]), B = 15000)
  # 
  omega.guess <- max(logit(2/pi*asin(cor(y1,y2))),0.05)
  cores=detectCores()
  cl <- makeCluster(cores[1]-1) #not to overload your computer
  registerDoParallel(cl)
  storeMATRIX <-  foreach(l=1:10000, .combine=cbind) %dopar% {
    
    library(copula)
    library(pracma)
    library(mvtnorm)
    MCMCobj.cop.correct.cut1 <- NewMCMCcut1.LogNormal.Gamma.T(y = cbind(y1,y2), mu.val = MCMCobj.marginal1.correct.cut1$mu.vec[5000+l], v.val = MCMCobj.marginal1.correct.cut1$v.vec[5000+l], a.val = MCMCobj.marginal2.correct.cut1$a.vec[5000+l], b.val = MCMCobj.marginal2.correct.cut1$b.vec[5000+l], omega.init = omega.guess, B = 500)
    c(MCMCobj.cop.correct.cut1$tau.vec[500])
    
  }
  stopCluster(cl)
  theta.correct.cut1[seed.num,] <- c(mean(MCMCobj.marginal1.correct.cut1$mu.vec[5001:15000]), mean(exp(MCMCobj.marginal1.correct.cut1$v.vec[5001:15000])), mean(exp(MCMCobj.marginal2.correct.cut1$a.vec[5001:15000])), mean(exp(MCMCobj.marginal2.correct.cut1$b.vec[5001:15000])), mean(storeMATRIX[1,]))
  theta.var.correct.cut1[seed.num,] <- c(var(MCMCobj.marginal1.correct.cut1$mu.vec[5001:15000]), var(exp(MCMCobj.marginal1.correct.cut1$v.vec[5001:15000])), var(exp(MCMCobj.marginal2.correct.cut1$a.vec[5001:15000])), var(exp(MCMCobj.marginal2.correct.cut1$b.vec[5001:15000])), var(storeMATRIX[1,]))
  lower.correct.cut1[seed.num,] <- c(quantile(MCMCobj.marginal1.correct.cut1$mu.vec[5001:15000],0.025), quantile(exp(MCMCobj.marginal1.correct.cut1$v.vec[5001:15000]), 0.025), quantile(exp(MCMCobj.marginal2.correct.cut1$a.vec[5001:15000]), 0.025), quantile(exp(MCMCobj.marginal2.correct.cut1$b.vec[5001:15000]), 0.025), quantile(storeMATRIX[1,], 0.025))
  upper.correct.cut1[seed.num,] <- c(quantile(MCMCobj.marginal1.correct.cut1$mu.vec[5001:15000],0.975), quantile(exp(MCMCobj.marginal1.correct.cut1$v.vec[5001:15000]), 0.975), quantile(exp(MCMCobj.marginal2.correct.cut1$a.vec[5001:15000]), 0.975), quantile(exp(MCMCobj.marginal2.correct.cut1$b.vec[5001:15000]), 0.975), quantile(storeMATRIX[1,], 0.975))
  cat("Completed seed number:", seed.num, "\n")
  
}
Bias.correct.cut1 <- rowMeans(apply(theta.correct.cut1,1,function(s){ s- true.val }))
RMSE.correct.cut1 <- sqrt(colMeans(t(apply(theta.correct.cut1,1,function(s){ (s - true.val)^2 })) ))
Coverage.correct.cut1 <- rowMeans(apply(lower.correct.cut1,1,function(s){ s <= true.val })*apply(upper.correct.cut1,1,function(s){ s >= true.val }))
KL.marginal1.correct.cut1 <- 0.5*log(1) - 0.5*log(theta.correct.cut1[,2]) + 0.5*(theta.correct.cut1[,2]/1) + 0.5*(theta.correct.cut1[,1]-1)^2 - 0.5
KL.marginal2.correct.cut1 <- rep(0,500)
KS.cop.vec <- rep(0,500)
u1.values <- seq(0.01,0.99,0.01)
u2.values <- u1.values
for(rs in 1:500){
  
  KL.marginal2.correct.cut1[rs] <- KL.GAMMA(shape1 = theta.correct.cut1[rs,3], rate1 = theta.correct.cut1[rs,4], shape2 = true.val[3], rate2 = true.val[4])
  KLGumbelCop <- function(u1, u2){
    
    a1 <- LogPDFTCop(F1 = u1, F2 = u2, tau = theta.correct.cut1[rs,5])
    a2 <- LogPDFTCop(F1 = u1, F2 = u2, tau = true.val[5])
    return(exp(a1)*(a1-a2))
    
  }
  grid.temp <- outer(u1.values,u2.values, FUN=KLGumbelCop)
  KS.cop.vec[rs] <- max(0,0.25*0.01^2*(sum(grid.temp[1:98,1:98]) + sum(grid.temp[1:98,2:99]) + sum(grid.temp[2:99,1:98]) + sum(grid.temp[2:99,2:99])))
  
}
Bias.correct.cut1
RMSE.correct.cut1
Coverage.correct.cut1
c(mean(KL.marginal2.correct.cut1),mean(KL.marginal2.correct.cut1),mean(KS.cop.vec))
```

#VB inference with correctly specified marginals and misspecified copula
```{r}
for(seed.num in 1:500){
  
  set.seed(seed.num)
  U <- rCopula(N, copula = cc)
  y1 <- qlnorm(U[,1], meanlog = 1, sdlog = 1)
  y2 <- qgamma(U[,2], shape=7, rate = 3)
  MLE <-univariateML::mlgamma(y2)
  VBobj.uncut <- VB.LogNormal.Gamma.Gumbel.uncut(y = cbind(y1,y2), m.init = c(mean(log(y1)), log(var(log(y1))), log(MLE[1]), log(MLE[2]), max(logit(2/pi*asin(cor(y1,y2))),0.05) ), C.init = diag(0.02,5))
  VBMainUncutMean <- colMeans(VBobj.uncut$mstore[5001:10000,])
  VBMainUncutVCov <- matrix(0,nrow=5,ncol=5)
  for(b in 5001:10000){
    
    VBMainUncutVCov <- VBMainUncutVCov + VBobj.uncut$Cstore[,,b] %*% t(VBobj.uncut$Cstore[,,b])
    
  }
  VBMainUncutVCov <- VBMainUncutVCov/5000
  theta.VB.uncut[seed.num,1] <- VBMainUncutMean[1]
  theta.VB.uncut[seed.num,2:4] <- exp(VBMainUncutMean[2:4] + 0.5*(diag(VBMainUncutVCov))[2:4])
  omega.post.draws <- rnorm(100000,mean=VBMainUncutMean[5],sd=sqrt(VBMainUncutVCov[5,5]))
  tau.post.draws <- exp(log(exp(exp(omega.post.draws)) - 1) - log(exp(exp(omega.post.draws)) + 1))
  theta.VB.uncut[seed.num,5] <- mean(tau.post.draws)
  theta.var.VB.uncut[seed.num,1] <- (diag(VBMainUncutVCov))[1]
  theta.var.VB.uncut[seed.num,2:4] <- (exp((diag(VBMainUncutVCov))[2:4])-1)*exp(2* VBMainUncutMean[2:4] + (diag(VBMainUncutVCov))[2:4])
  theta.var.VB.uncut[seed.num,5] <- var(tau.post.draws)
  lower.VB.uncut[seed.num,1] <- qnorm(0.025,mean = VBMainUncutMean[1], sd = sqrt((diag(VBMainUncutVCov))[1]) )
  lower.VB.uncut[seed.num,2:4] <- qlnorm(0.025,meanlog = VBMainUncutMean[2:4], sdlog = sqrt((diag(VBMainUncutVCov))[2:4]))
  lower.VB.uncut[seed.num,5] <- quantile(tau.post.draws,0.025)
  upper.VB.uncut[seed.num,1] <- qnorm(0.975,mean = VBMainUncutMean[1], sd = sqrt((diag(VBMainUncutVCov))[1]) )
  upper.VB.uncut[seed.num,2:4] <- qlnorm(0.975,meanlog = VBMainUncutMean[2:4], sdlog = sqrt((diag(VBMainUncutVCov))[2:4]))
  upper.VB.uncut[seed.num,5] <- quantile(tau.post.draws,0.975)
  cat("Completed seed number:", seed.num, "\n")
  
}
Bias.VB.uncut <- rowMeans(apply(theta.VB.uncut,1,function(s){ s- true.val }))
RMSE.VB.uncut <- sqrt(colMeans(t(apply(theta.VB.uncut,1,function(s){ (s - true.val)^2 })) ))
Coverage.VB.uncut <- rowMeans(apply(lower.VB.uncut,1,function(s){ s <= true.val })*apply(upper.VB.uncut,1,function(s){ s >= true.val }))
KL.marginal1.VB.uncut <- 0.5*log(1) - 0.5*log(theta.VB.uncut[,2]) + 0.5*(theta.VB.uncut[,2]/1) + 0.5*(theta.VB.uncut[,1]-1)^2 - 0.5
KL.marginal2.VB.uncut <- rep(0,500)
KS.cop.vec <- rep(0,500)
u1.values <- seq(0.01,0.99,0.01)
u2.values <- u1.values
for(rs in 1:500){
  
  KL.marginal2.VB.uncut[rs] <- KL.GAMMA(shape1 = theta.VB.uncut[rs,3], rate1 = theta.VB.uncut[rs,4], shape2 = true.val[3], rate2 = true.val[4])
  KLGumbelCop <- function(u1, u2){
    
    a1 <- LogPDFGumbelCop(F1 = u1, F2 = u2, tau = theta.VB.uncut[rs,5])
    a2 <- LogPDFTCop(F1 = u1, F2 = u2, tau = true.val[5])
    return(exp(a1)*(a1-a2))
    
  }
  grid.temp <- outer(u1.values,u2.values, FUN=KLGumbelCop)
  KS.cop.vec[rs] <- max(0,0.25*0.01^2*(sum(grid.temp[1:98,1:98]) + sum(grid.temp[1:98,2:99]) + sum(grid.temp[2:99,1:98]) + sum(grid.temp[2:99,2:99])))
  
}
Bias.VB.uncut
RMSE.VB.uncut
Coverage.VB.uncut
c(mean(KL.marginal1.VB.uncut),mean(KL.marginal2.VB.uncut),mean(KS.cop.vec))
```

#VB cut I posterior inference with correctly specified marginals and misspecified copula
```{r}
for(seed.num in 1:500){
  
  set.seed(seed.num)
  U <- rCopula(N, copula = cc)
  y1 <- qlnorm(U[,1], meanlog = 1, sdlog = 1)
  y2 <- qgamma(U[,2], shape=7, rate = 3)
  MLE <-univariateML::mlgamma(y2)
  omega.guess <- max(logit(2/pi*asin(cor(y1,y2))),0.05)
  VBobj.marginal1.cut1 <- VB.Marginal.LogNormal.cut1(y = y1, m.init = c( mean(log(y1)), log(var(log(y1))) ), C.init = diag(0.1,2))
  VBobj.marginal2.cut1 <- VB.Marginal.Gamma.cut1(y = y2, m.init = c(MLE[1],MLE[2]), C.init = diag(0.1,2))
  VBMainCutMean <- c(colMeans(VBobj.marginal1.cut1$mstore[5001:10000,]),colMeans(VBobj.marginal2.cut1$mstore[5001:10000,]),omega.guess)
  VBMainCutCov <- matrix(0,nrow=5,ncol=5)
  for(b in 5001:10000){
    
    VBMainCutCov[1:2,1:2] <- VBMainCutCov[1:2,1:2] + VBobj.marginal1.cut1$Cstore[,,b] %*% t(VBobj.marginal1.cut1$Cstore[,,b])
    VBMainCutCov[3:4,3:4] <- VBMainCutCov[3:4,3:4] + VBobj.marginal2.cut1$Cstore[,,b] %*% t(VBobj.marginal2.cut1$Cstore[,,b])
    
  }
  VBMainCutCov <- VBMainCutCov/5000
  VBMainCutCov[1:4,1:4] <- t(chol(VBMainCutCov[1:4,1:4]))
  VBMainCutCov[5,5] <- 0.1
  VBobj.cop.cut1 <- VB.LogNormal.Gamma.Gumbel.cut1(y = cbind(y1, y2), m.init = VBMainCutMean, C.init = VBMainCutCov)
  VBMainCutCov <- matrix(0,nrow=5,ncol=5)
  for(b in 5001:10000){
    
    VBMainCutCov <- VBMainCutCov + VBobj.cop.cut1$Cstore[,,b] %*% t(VBobj.cop.cut1$Cstore[,,b])
    
  }
  VBMainCutCov <- VBMainCutCov/5000
  VBMainCutMean <- colMeans(VBobj.cop.cut1$mstore[5001:10000,])
  theta.VB.cut1[seed.num,1] <- VBMainCutMean[1]
  theta.VB.cut1[seed.num,2:4] <- exp(VBMainCutMean[2:4] + 0.5*(diag(VBMainCutCov))[2:4])
  omega.post.draws <- rnorm(100000,mean=VBMainCutMean[5],sd=sqrt(VBMainCutCov[5,5]))
  tau.post.draws <- exp(log(exp(exp(omega.post.draws)) - 1) - log(exp(exp(omega.post.draws)) + 1))
  theta.VB.cut1[seed.num,5] <- mean(tau.post.draws)
  theta.var.VB.cut1[seed.num,1] <- (diag(VBMainCutCov))[1]
  theta.var.VB.cut1[seed.num,2:4] <- (exp((diag(VBMainCutCov))[2:4])-1)*exp(2* VBMainCutMean[2:4] + (diag(VBMainCutCov))[2:4])
  theta.var.VB.cut1[seed.num,5] <- var(tau.post.draws)
  lower.VB.cut1[seed.num,1] <- qnorm(0.025,mean = VBMainCutMean[1], sd = sqrt((diag(VBMainCutCov))[1]) )
  lower.VB.cut1[seed.num,2:4] <- qlnorm(0.025,meanlog = VBMainCutMean[2:4], sdlog = sqrt((diag(VBMainCutCov))[2:4]))
  lower.VB.cut1[seed.num,5] <- quantile(tau.post.draws,0.025)
  upper.VB.cut1[seed.num,1] <- qnorm(0.975,mean = VBMainCutMean[1], sd = sqrt((diag(VBMainCutCov))[1]) )
  upper.VB.cut1[seed.num,2:4] <- qlnorm(0.975,meanlog = VBMainCutMean[2:4], sdlog = sqrt((diag(VBMainCutCov))[2:4]))
  upper.VB.cut1[seed.num,5] <- quantile(tau.post.draws,0.975)
  cat("Completed seed number:", seed.num, "\n")

  
}
Bias.VB.cut1 <- rowMeans(apply(theta.VB.cut1,1,function(s){ s- true.val }))
RMSE.VB.cut1 <- sqrt(colMeans(t(apply(theta.VB.cut1,1,function(s){ (s - true.val)^2 })) ))
Coverage.VB.cut1 <- rowMeans(apply(lower.VB.cut1,1,function(s){ s <= true.val })*apply(upper.VB.cut1,1,function(s){ s >= true.val }))
KL.marginal1.VB.cut1 <- 0.5*log(1) - 0.5*log(theta.VB.cut1[,2]) + 0.5*(theta.VB.cut1[,2]/1) + 0.5*(theta.VB.cut1[,1]-1)^2 - 0.5
KL.marginal2.VB.cut1 <- rep(0,500)
KS.cop.vec <- rep(0,500)
u1.values <- seq(0.01,0.99,0.01)
u2.values <- u1.values
for(rs in 1:500){
  
  KL.marginal2.VB.cut1[rs] <- KL.GAMMA(shape1 = theta.VB.cut1[rs,3], rate1 = theta.VB.cut1[rs,4], shape2 = true.val[3], rate2 = true.val[4])
  KLGumbelCop <- function(u1, u2){
    
    a1 <- LogPDFGumbelCop(F1 = u1, F2 = u2, tau = theta.VB.cut1[rs,5])
    a2 <- LogPDFTCop(F1 = u1, F2 = u2, tau = true.val[5])
    return(exp(a1)*(a1-a2))
    
  }
  grid.temp <- outer(u1.values,u2.values, FUN=KLGumbelCop)
  KS.cop.vec[rs] <- max(0,0.25*0.01^2*(sum(grid.temp[1:98,1:98]) + sum(grid.temp[1:98,2:99]) + sum(grid.temp[2:99,1:98]) + sum(grid.temp[2:99,2:99])))
  
}
Bias.VB.cut1
RMSE.VB.cut1
Coverage.VB.cut1
c(mean(KL.marginal1.VB.cut1),mean(KL.marginal2.VB.cut1),mean(KS.cop.vec))
```


