---
title: "Implementation of Cut II Posterior for Macroeconomic dataset"
author: "CutPosteriorAuthors"
date: "2024-05-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load required libraries
```{r}
library(sn)
library(mvtnorm)
library(pracma)
library(linpk)
library(Rcpp)
library(RcppArmadillo)
library(Matrix)
library(sm)
library(ks)
```

Load customised functions for analysis
```{r}
sourceCpp("NewUSMacroSupportFile.cpp")
source("NewUSMacroSupportFile.R")
```

Read dataset
```{r}
dat <- read.csv("macro2022Q3.csv",header=TRUE)
y1 <- dat$GDP.Growth
y2 <- dat$Inflation
y3 <- dat$IR
y4 <- dat$UR
y1 <- y1[1:274]; y2 <- y2[1:274]; y3 <- y3[1:274]; y4 <- y4[1:274]
```

Read initial guess for marginal parameters
```{r}
MarginalPar <- scan("TrialMarginal.txt")
u1 <- pst(x = y1,xi = MarginalPar[1], omega = exp(MarginalPar[2]), alpha = MarginalPar[3], nu = exp(MarginalPar[4]))
u2 <- pst(x = y2,xi = MarginalPar[5], omega = exp(MarginalPar[6]), alpha = MarginalPar[7], nu = exp(MarginalPar[8]))
u3 <- (pst(x = y3,xi = MarginalPar[9], omega = exp(MarginalPar[10]), alpha = MarginalPar[11], nu = exp(MarginalPar[12])) - pst(x = 0, xi = MarginalPar[9], omega = exp(MarginalPar[10]), alpha = MarginalPar[11], nu = exp(MarginalPar[12]))) /(1 - pst(x = 0, xi = MarginalPar[9], omega = exp(MarginalPar[10]), alpha = MarginalPar[11], nu = exp(MarginalPar[12])))
u4 <- (pst(x = y4,xi = MarginalPar[13], omega = exp(MarginalPar[14]), alpha = MarginalPar[15], nu = exp(MarginalPar[16])) - pst(x = 0, xi = MarginalPar[13], omega = exp(MarginalPar[14]), alpha = MarginalPar[15], nu = exp(MarginalPar[16]))) /(1 - pst(x = 0, xi = MarginalPar[13], omega = exp(MarginalPar[14]), alpha = MarginalPar[15], nu = exp(MarginalPar[16])))
```

Use entire dataset to compute MLE for model parmaeters
```{r}
UMAT <- cbind(u1,u2,u3,u4)
OmegaArray.initial <- array(0,dim=c(4,4,5))
OmegaArray.initial[,,1] <- cor(qnorm(UMAT))
for(i in 1:4){
  for(j in 1:4){
    
    OmegaArray.initial[i,j,2] <- cor(qnorm(UMAT[1:(nrow(UMAT)-1),i]),qnorm(UMAT[2:(nrow(UMAT)),j]))
    
  }
}
for(i in 1:4){
  for(j in 1:4){
    
    OmegaArray.initial[i,j,3] <- cor(qnorm(UMAT[1:(nrow(UMAT)-2),i]),qnorm(UMAT[3:(nrow(UMAT)),j]))
    
  }
}
for(i in 1:4){
  for(j in 1:4){
    
    OmegaArray.initial[i,j,4] <- cor(qnorm(UMAT[1:(nrow(UMAT)-3),i]),qnorm(UMAT[4:(nrow(UMAT)),j]))
    
  }
}
for(i in 1:4){
  for(j in 1:4){
    
    OmegaArray.initial[i,j,5] <- cor(qnorm(UMAT[1:(nrow(UMAT)-4),i]),qnorm(UMAT[5:(nrow(UMAT)),j]))
    
  }
}

BigT <- 4*length(y1)
m <- ncol(UMAT)
BigOmega <- matrix(0,nrow=BigT,ncol=BigT)
for(s in 1:5){
  for(t in max(1,s-4):min(s+4,nrow(UMAT))){
    
    a.t <- (t-1)*m+1
    b.t <- m*t
    a.s <- (s-1)*m+1
    b.s <- s*m
    
    if(s<t){
      
      BigOmega[a.s:b.s,a.t:b.t] <- OmegaArray.initial[,,abs(t-s)+1]
      
    }
    if(s>t){
      
      BigOmega[a.s:b.s,a.t:b.t] <- t(OmegaArray.initial[,,abs(t-s)+1])
      
    }
    if(s==t){
      
      BigOmega[a.s:b.s,a.t:b.t] <- OmegaArray.initial[,,1]
      
    }
  }
}
BigOmega2 <- BigOmega[1:20,1:20]
PhiBlock <- ConvertSemiPartialCorCpp(BigOmega2) 
Phi0 <- PhiBlock[1:4,1:4]
Phi1 <- PhiBlock[5:8,1:4]
Phi2 <- PhiBlock[9:12,1:4]
Phi3 <- PhiBlock[13:16,1:4]
Phi4 <- PhiBlock[17:20,1:4]
PhiArray.MLE <- array(0, dim=c(4,4,5))
PhiArray.MLE[,,1] <- Phi0
PhiArray.MLE[,,2] <- Phi1
PhiArray.MLE[,,3] <- Phi2
PhiArray.MLE[,,4] <- Phi3
PhiArray.MLE[,,5] <- Phi4
expandedPhi <- qnorm(0.5*(c(PhiArray.MLE[2:4,1,1],PhiArray.MLE[3:4,2,1], PhiArray.MLE[4,3,1], PhiArray.MLE[,,2:5])+1))
```

Compute discrete-valued standardised rank of entire dataset
```{r}
a1 <- (rank(y1)-1)/(length(y1)+1)
a2 <- (rank(y2)-1)/(length(y2)+1)
a3 <- (rank(y3)-1)/(length(y3)+1)
a4 <- (rank(y4)-1)/(length(y4)+1)
b1 <- rank(y1)/(length(y1)+1)
b2 <- rank(y2)/(length(y2)+1)
b3 <- rank(y3)/(length(y3)+1)
b4 <- rank(y4)/(length(y4)+1)
Amat.rank <- cbind(a1,a2,a3,a4)
Bmat.rank <- cbind(b1,b2,b3,b4)
```

Compute Cut I posteriors of model parameters for initialisation of Cut II posteriors (based on entire dataset)
```{r}
source("InitialisationCutII.R")
```

Compute Cut II posterior of copula parameters (based on entire dataset)
```{r}
etaMatTry <- matrix(rt(length(y1)*4,df=3),nrow=length(y1),ncol=4)
wMatTry <- log(matrix(1/rgamma(length(y1)*4,shape=2,rate=4),nrow=length(y1),ncol=4))
ReverseVBCopObj <- NewCopulaVBReverseCutRversion(AMAT = Amat.rank, BMAT = Bmat.rank, muInit = VB.uncut.obj$muMAT[10000,], CInit = VB.uncut.obj$CArray[,,10000], EtaMatInit = etaMatTry, WMatInit = wMatTry, maxRuns = 1000)
```

Compute Cut II posterior of marginal parameters (based on entire dataset)
```{r}
muInit.marginal <- rep(0,91)
muInit.marginal[1:75] <- ReverseVBCopObj$MuStore[1000,1:75]
CInit.marginal <- diag(0.1,91)
CInit.marginal[1:75,1:75] <- ReverseVBCopObj$CStore[,,1000]
ReverseVBMargObj <- NewCopulaVBReverseCutMarginalPar(y1 = y1, y2=y2, y3=y3, y4=y4, maxRuns = 10000, mu.init = muInit.marginal, C.init = CInit.marginal)
ParDraws <- rmvnorm(50000,mean = ReverseVBMargObj$muMAT[10000,], sigma = (ReverseVBMargObj$CArray[,,10000] %*% t(ReverseVBMargObj$CArray[,,10000]) ) )
ParDraws <- ParDraws[,c(76:91,1:75)]
```

Compute 4-step forward log-score metric
```{r}
MarginalDrawsKernelEst <- matrix(0,nrow=32,ncol=183)
for(tprime in 84:266){
  
  y1.temp <- y1[1:tprime]
  y2.temp <- y2[1:tprime]
  y3.temp <- y3[1:tprime]
  y4.temp <- y4[1:tprime]
  a1.temp <- (rank(y1.temp)-1)/(length(y1.temp)+1)
  a2.temp <- (rank(y2.temp)-1)/(length(y2.temp)+1)
  a3.temp <- (rank(y3.temp)-1)/(length(y3.temp)+1)
  a4.temp <- (rank(y4.temp)-1)/(length(y4.temp)+1)
  b1.temp <- rank(y1.temp)/(length(y1.temp)+1)
  b2.temp <- rank(y2.temp)/(length(y2.temp)+1)
  b3.temp <- rank(y3.temp)/(length(y3.temp)+1)
  b4.temp <- rank(y4.temp)/(length(y4.temp)+1)
  Amat.rank.temp <- cbind(a1.temp,a2.temp,a3.temp,a4.temp)
  Bmat.rank.temp <- cbind(b1.temp,b2.temp,b3.temp,b4.temp)
  etaMatTry <- matrix(rt(length(y1.temp)*4,df=3),nrow=length(y1.temp),ncol=4)
  wMatTry <- log(matrix(1/rgamma(length(y1.temp)*4,shape=2,rate=4),nrow=length(y1.temp),ncol=4))
  ReverseVBCopObj.temp <- NewCopulaVBReverseCutRversion(AMAT = Amat.rank.temp, BMAT = Bmat.rank.temp, muInit = VB.uncut.obj$muMAT[10000,], CInit = VB.uncut.obj$CArray[,,10000], EtaMatInit = etaMatTry, WMatInit = wMatTry, maxRuns = 1000)
  muInit.marginal <- rep(0,91)
  muInit.marginal[1:75] <- ReverseVBCopObj.temp$MuStore[1000,1:75]
  CInit.marginal <- diag(0.1,91)
  CInit.marginal[1:75,1:75] <- ReverseVBCopObj.temp$CStore[,,1000]
  ReverseVBMargObj.temp <- NewCopulaVBReverseCutMarginalPar(y1 = y1.temp, y2=y2.temp, y3=y3.temp, y4=y4.temp, maxRuns = 10000, mu.init = muInit.marginal, C.init = CInit.marginal)
  ParDraws <- rmvnorm(50000,mean = ReverseVBMargObj$muMAT[10000,], sigma = (ReverseVBMargObj$CArray[,,10000] %*% t(ReverseVBMargObj$CArray[,,10000]) ) )
  ParDraws <- ParDraws[,c(76:91,1:75)]
  
  
  u1.draw.MAT <- matrix(0,nrow=(tprime+8),ncol=5000)
  u2.draw.MAT <- matrix(0,nrow=(tprime+8),ncol=5000)
  u3.draw.MAT <- matrix(0,nrow=(tprime+8),ncol=5000)
  u4.draw.MAT <- matrix(0,nrow=(tprime+8),ncol=5000)
  ParDraws.temp <- rmvnorm(5000,mean = ReverseVBMargObj.temp[10000,], sigma = (ReverseVBMargObj.temp$CArray[,,10000] %*% t(ReverseVBMargObj.temp$CArray[,,10000]) ) )
  ParDraws.temp <- ParDraws.temp[,c(76:91,1:75)]
  Whistory.All.MAT <- matrix(0,nrow=5000,ncol=(4*(tprime+8)))
  for(l in 1:5000){
    
    u1.draw.MAT[,l] <- pst(y1[1:(tprime+8)], xi = ParDraws.temp[l,1], omega = exp(ParDraws.temp[l,2]), alpha = ParDraws.temp[l,3], nu = exp(ParDraws.temp[l,4]))
    u2.draw.MAT[,l] <- pst(y2[1:(tprime+8)], xi = ParDraws.temp[l,5], omega = exp(ParDraws.temp[l,6]), alpha = ParDraws.temp[l,7], nu = exp(ParDraws.temp[l,8]))
    u3.draw.MAT[,l] <- (pst(y3[1:(tprime+8)], xi = ParDraws.temp[l,9], omega = exp(ParDraws.temp[l,10]), alpha = ParDraws.temp[l,11], nu = exp(ParDraws.temp[l,12])) - pst(0, xi = ParDraws.temp[l,9], omega = exp(ParDraws.temp[l,10]), alpha = ParDraws.temp[l,11], nu = exp(ParDraws.temp[l,12])))/(1 - pst(0, xi = ParDraws.temp[l,9], omega = exp(ParDraws.temp[l,10]), alpha = ParDraws.temp[l,11], nu = exp(ParDraws.temp[l,12])))
    u4.draw.MAT[,l] <- (pst(y4[1:(tprime+8)], xi = ParDraws.temp[l,13], omega = exp(ParDraws.temp[l,14]), alpha = ParDraws.temp[l,15], nu = exp(ParDraws.temp[l,16])) - pst(0, xi = ParDraws.temp[l,13], omega = exp(ParDraws.temp[l,14]), alpha = ParDraws.temp[l,15], nu = exp(ParDraws.temp[l,16])))/(1 - pst(0, xi = ParDraws.temp[l,13], omega = exp(ParDraws.temp[l,14]), alpha = ParDraws.temp[l,15], nu = exp(ParDraws.temp[l,16])))
    Whistory.All.MAT[l,] <- c(rbind(rev(u1.draw.MAT[,l]),rev(u2.draw.MAT[,l]),rev(u3.draw.MAT[,l]), rev(u4.draw.MAT[,l]) ))
    
  }
  
  Whistory.All.MAT <- qnorm(Whistory.All.MAT)
  DrawsMAT <- pnorm(PredictiveDrawRun(ParDraws = ParDraws.temp[,17:86], WhistoryMAT = Whistory.All.MAT))
  MarginalDraws <- matrix(0,nrow=32,ncol=5000)
  DrawsMAT <- pmax(pmin(DrawsMAT,matrix(0.98,nrow=32,ncol=5000)), matrix(0.02,nrow=32,ncol=5000))
  for(l in 1:5000){
    
    MarginalDraws[seq(1,32,4),l] <- qst(DrawsMAT[seq(1,32,4),l], xi = ParDraws.temp[l,1], omega =  exp(ParDraws.temp[l,2]), alpha = ParDraws.temp[l,3] , nu = exp(ParDraws.temp[l,4]) )
    MarginalDraws[seq(2,32,4),l] <- qst(DrawsMAT[seq(2,32,4),l], xi = ParDraws.temp[l,5], omega =  exp(ParDraws.temp[l,6]), alpha = ParDraws.temp[l,7] , nu = exp(ParDraws.temp[l,8]) )
    F03.temp <- pst(0, xi = ParDraws.temp[l,9], omega =  exp(ParDraws.temp[l,10]), alpha = ParDraws.temp[l,11], nu = exp(ParDraws.temp[l,12]) )
    Draw3 <- (DrawsMAT[seq(3,32,4),l]*(1-F03.temp)+F03.temp)
    Draw3 <- pmax(pmin(Draw3,rep(0.95,8)),rep(0.05,8))
    MarginalDraws[seq(3,32,4),l] <- qst(Draw3, xi = ParDraws.temp[l,9], omega =  exp(ParDraws.temp[l,10]), alpha = ParDraws.temp[l,11], nu = exp(ParDraws.temp[l,12]) )
    F04.temp <- pst(0, xi = ParDraws.temp[l,13], omega =  exp(ParDraws.temp[l,14]), alpha = ParDraws.temp[l,15], nu = exp(ParDraws.temp[l,16]) )
    Draw4 <- (DrawsMAT[seq(4,32,4),l]*(1-F04.temp)+F04.temp)
    Draw4 <- pmax(pmin(Draw4,rep(0.95,8)),rep(0.05,8))
    MarginalDraws[seq(4,32,4),l] <- qst(Draw4, xi = ParDraws.temp[l,13], omega =  exp(ParDraws.temp[l,14]), alpha = ParDraws.temp[l,15], nu = exp(ParDraws.temp[l,16]) )
    
  }
  for(j in seq(1,32,4)){
    
    MarginalDrawsKernelEst[j,(tprime-83)]  <- log(sm.density(x=MarginalDraws[j,], eval.points=c(y1[tprime+ceiling(j/4)]), display="none")$estimate)
    
  }
  for(j in seq(2,32,4)){
    
    MarginalDrawsKernelEst[j,(tprime-83)]  <- log(sm.density(x=MarginalDraws[j,], eval.points=c(y2[tprime+ceiling(j/4)]), display="none")$estimate)
    
  }
  for(j in seq(3,32,4)){
    
    MarginalDrawsKernelEst[j,(tprime-83)]  <- log(sm.density(x=MarginalDraws[j,], eval.points=c(y3[tprime+ceiling(j/4)]), display="none")$estimate)
    
  }
  for(j in seq(4,32,4)){
    
    MarginalDrawsKernelEst[j,(tprime-83)]  <- log(sm.density(x=MarginalDraws[j,], eval.points=c(y4[tprime+ceiling(j/4)]), display="none")$estimate)
    
  }
  
}
```